---
title: "Highcharter - Barplots"
description: |
  Highcharter is one of the best interactive visualization package available in R. Barplots are the most commonly used plots in the data visualization world. 
author:
  - name: Umesh Narayanappa
date: 09-12-2020
output: distill::distill_article
editor_options: 
  chunk_output_type: console
---


## What is highcharter?

Highcharter is a R wrapper for Highcharts javascript library and its modules. Highcharts is very flexible and customizable javascript charting library and it has a great and powerful API.


### Data View

We are going to use this dataset for plotting different variations of the barplots

```{r}

library(tidyverse)
library(highcharter)
library(DT)

DT::datatable(mpg)

```


### Default Bar Plot is a Horizontal plot

```{r, layout="l-body-outset"}

makers <- count(mpg, manufacturer)

hchart(
  makers,
  "bar",
  hcaes(x=manufacturer, y= n)
)

```


### How to make it a vertical bar plot

```{r, layout="l-body-outset"}

hchart(
  makers,
  "column",
  hcaes(x=manufacturer, y= n)
)

```

### How to change the default color of the bars

```{r, layout="l-body-outset"}

hchart(
  makers,
  "column",
  hcaes(x=manufacturer, y= n),
  color = "maroon"
)

```

### How to sort the bars in the plot in a order

```{r, layout="l-body-outset"}

mpg %>%
  group_by(manufacturer) %>%
  count() %>%
  arrange(desc(n)) -> makers 

hchart(
  makers,
  "column",
  hcaes(x=manufacturer, y= n),
  color = "maroon"
)

```


### How to Create Side by Side Bar plot

```{r, layout="l-body-outset"}

mpgman2 <- count(mpg, manufacturer, year)

hchart(
  mpgman2, 
  "bar",
  hcaes(x = manufacturer, y = n, group = year),
  color = c("maroon", "darkgrey"),
  name = c("Year 1999", "Year 2008")
  )

```


### How to Create Side by Side Column plot

```{r, layout="l-body-outset"}

mpgman2 <- count(mpg, manufacturer, year)

hchart(
  mpgman2, 
  "column",
  hcaes(x = manufacturer, y = n, group = year),
  color = c("maroon", "darkgrey"),
  name = c("Year 1999", "Year 2008")
  )

```


### How to Sort the Bars based on Categorical Attribute


```{r, layout="l-body-outset"}

mpg %>%
  group_by(manufacturer, year) %>%
  count() %>%
  arrange(desc(manufacturer)) -> makers 

hchart(
  makers, 
  "column",
  hcaes(x = manufacturer, y = n, group = year),
  color = c("maroon", "darkgrey"),
  name = c("Year 1999", "Year 2008")
  )

```


### How to Create Stacked Bars plot


```{r, layout="l-body-outset"}

mpgman2 <- count(mpg, manufacturer, year)

hchart(
  mpgman2, 
  "column",
  hcaes(x = manufacturer, y = n, group = year),
  color = c("maroon", "darkgrey"),
  name = c("Year 1999", "Year 2008"),
  stacking = "normal"
  )

```


## Credits

* Thanks to Joshua Kunst for creating such a wonderful package for visualization
* Thanks to the whole team at RStudio and the whole R community 