---
title: "Highcharter - Line Plots"
description: |
  Line plots a very common type of plots specially in finance domain. These plots are useful for viewing the variation in a numerical attribute over a time period.
author:
  - name: Umesh Narayanappa
date: 11-12-2020
output: distill::distill_article
editor_options: 
  chunk_output_type: console
---


## What is highcharter?

Highcharter is a R wrapper for Highcharts javascript library and its modules. Highcharts is very flexible and customizable javascript charting library and it has a great and powerful API.


### Line plot

Line plots a very common type of plots specially in finance domain. These plots are useful for viewing the variation in a numerical attribute over a time period.

### Data View

We are going to use this dataset for plotting different variations of the scatterplots

```{r}

library(tidyverse)
library(highcharter)
library(DT)

data(economics_long, package = "ggplot2")

economics_long2 <- dplyr::filter(economics_long, variable %in% c("pop", "uempmed", "unemploy"))

DT::datatable(economics_long2)

```


### Default Line Plot

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("unemploy")),
  "line",
  hcaes(x=date,
        y=value01)
)

```

### Changing the color of the line

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("unemploy")),
  "line",
  hcaes(x=date,
        y=value01),
  color = "#fca311"
)

```


### Plotting two numerical variables over time

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("pop","unemploy")),
  "line",
  hcaes(x=date,
        y=value01,
        group = variable),
  color = c("#00296b","#fca311")
)

```



### Plotting three numerical variables over time

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("pop","unemploy", "uempmed")),
  "line",
  hcaes(x=date,
        y=value01,
        group = variable),
  color = c("#00296b","#fca311","#177e89")
)

```



### Legend - Horizontal Bottom Right 

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("pop","unemploy", "uempmed")),
  "line",
  hcaes(x=date,
        y=value01,
        group = variable),
  color = c("#00296b","#fca311","#177e89")
) %>%
  hc_legend(
    layout = "horizontal",
    align = "right",
    verticalAlign = "bottom"
    
  )

```


### Legend - Horizontal Bottom Right 

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("pop","unemploy", "uempmed")),
  "line",
  hcaes(x=date,
        y=value01,
        group = variable),
  color = c("#00296b","#fca311","#177e89")
) %>%
  hc_legend(
    layout = "horizontal",
    align = "left",
    verticalAlign = "bottom"
    
  )

```


### Legend - Horizontal Top Right 

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("pop","unemploy", "uempmed")),
  "line",
  hcaes(x=date,
        y=value01,
        group = variable),
  color = c("#00296b","#00509d","#fdc500")
) %>%
  hc_legend(
    layout = "horizontal",
    align = "right",
    verticalAlign = "top"
    
  )

```

### Legend - Horizontal Top Left 

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("pop","unemploy", "uempmed")),
  "line",
  hcaes(x=date,
        y=value01,
        group = variable),
  color = c("#00296b","#00509d","#fdc500")
) %>%
  hc_legend(
    layout = "horizontal",
    align = "left",
    verticalAlign = "top"
    
  )

```


### Legend - Vertical Top Right 

```{r}

hchart(
  dplyr::filter(economics_long, variable %in% c("pop","unemploy", "uempmed")),
  "line",
  hcaes(x=date,
        y=value01,
        group = variable),
  color = c("#00296b","#00509d","#fdc500")
) %>%
  hc_legend(
    layout = "vertical",
    align = "right",
    verticalAlign = "top"
    
  )

```


### Plotting more numerical variables over time

```{r}

hchart(
  economics_long,
  "line",
  hcaes(x=date,
        y=value01,
        group = variable),
  color = c("#00296b","#fca311","#177e89", "#8f2d56", "#6d597a")
)

```


## Credits

* Thanks to Joshua Kunst for creating such a wonderful package for visualization
* Thanks to the whole team at RStudio and the whole R community 